---
title: "Smart Typst Tables Example"
format:
  typst:
    toc: false
    papersize: us-letter
    margin:
      x: 0.75in
      y: 0.8in
filters:
  - ../_extensions/smart-typst-tables/smart-typst-tables.lua
execute:
  echo: false
  warning: true
---

This document demonstrates several table shapes handled by
`smart-typst-tables`: wide transaction data, compact tables, numeric summaries,
long text, and explicit overrides when a course handout needs a specific layout.

```{r}
#| label: setup
source_path <- c(
  "R/smart_table.R",
  "../R/smart_table.R",
  "_extensions/smart-typst-tables/smart_table.R",
  "../_extensions/smart-typst-tables/smart_table.R"
)
source(source_path[file.exists(source_path)][1])

transactions <- read.csv(
  if (file.exists("examples/data/transactions.csv")) {
    "examples/data/transactions.csv"
  } else {
    "data/transactions.csv"
  },
  check.names = FALSE
)
```

## 1. Default academic table

The default profile computes compact fixed widths for identifiers, dates, and
currency columns, then gives the remaining width to the descriptive text column.

```{r}
#| output: asis
transactions |>
  smart_table(
    caption = "Transactions d'assurance — profil académique",
    profile = "academic",
    width = "100%",
    auto_widths = TRUE,
    striped = TRUE
  )
```

## 2. Compact table with notes

The compact profile is useful for dense course notes and appendices.

```{r}
#| output: asis
transactions |>
  smart_table(
    caption = "Même table, profil compact",
    profile = "compact",
    width = "100%",
    striped = TRUE,
    notes = "Note : les tirets indiquent une prime non applicable à la transaction."
  )
```

## 3. Numeric summary table

Numeric, percentage, and currency-like columns are right-aligned and kept narrow.

```{r}
#| output: asis
summary_table <- data.frame(
  Segment = c("Nouvelles ventes", "Renouvellements", "Annulations", "Modifications"),
  Polices = c(128, 94, 17, 31),
  `Prime moyenne` = c("1 240 $", "980 $", "-", "610 $"),
  `Ratio de conversion` = c("42.5%", "71.0%", "8.2%", "19.4%"),
  `Commentaire pédagogique` = c(
    "Fort volume mais prime moyenne élevée.",
    "Portefeuille stable avec bonne rétention.",
    "À surveiller dans les exercices de sensibilité.",
    "Effet mixte sur la prime totale."
  ),
  check.names = FALSE
)

summary_table |>
  smart_table(
    caption = "Résumé numérique avec pourcentages",
    profile = "teaching",
    width = "100%",
    striped = TRUE
  )
```

## 4. Long-text teaching table

Long text columns absorb the available width. Short categorical and numeric
columns stay compact.

```{r}
#| output: asis
rubric <- data.frame(
  Critère = c("Définition", "Calcul", "Interprétation", "Communication"),
  Poids = c("20%", "30%", "35%", "15%"),
  `Attente pour une réponse complète` = c(
    "Définit clairement la prime annuelle, la prime du terme et la date de transaction.",
    "Montre les étapes de calcul et indique les hypothèses utilisées.",
    "Explique pourquoi une annulation ou une modification change le revenu reconnu.",
    "Présente les résultats avec unités, dates et libellés cohérents."
  ),
  check.names = FALSE
)

rubric |>
  smart_table(
    caption = "Grille d'évaluation courte",
    profile = "spacious",
    width = "100%",
    striped = FALSE,
    notes = "Exemple conçu pour des notes de cours imprimées."
  )
```

## 5. Explicit overrides when needed

The extension avoids manual tuning for common cases, but still allows explicit
widths and preferred header breaks when a table has a known pedagogical layout.

```{r}
#| output: asis
comparison <- data.frame(
  Méthode = c("Typst auto", "Largeurs manuelles", "smart_table()"),
  `Temps de préparation` = c("Faible", "Élevé", "Faible"),
  `Qualité sur tables larges` = c("Variable", "Bonne", "Bonne par défaut"),
  `Quand l'utiliser` = c(
    "Petites tables simples.",
    "Table finale très contrôlée.",
    "Documents de cours avec plusieurs tableaux récurrents."
  ),
  check.names = FALSE
)

comparison |>
  smart_table(
    caption = "Comparaison des stratégies de mise en page",
    profile = "academic",
    auto_widths = FALSE,
    col_widths = c("1.2fr", "1fr", "1.2fr", "2.6fr"),
    header_breaks = list(
      "Temps de préparation" = c("Temps de", "préparation"),
      "Qualité sur tables larges" = c("Qualité sur", "tables larges"),
      "Quand l'utiliser" = c("Quand", "l'utiliser")
    ),
    striped = TRUE
  )
```
